Locality and Computational Reliability of Linear Response Calculations for Molecular Systems

### 2019-01-18 AnalyticStructure_main.tex
added the C60 panel in figure 6 and the supplementary file

From the above expressions we can see that, for molecular systems,the value $\epsilon=0$
represents a \emph{threshold}~\footnote{The threshold value in general should correspond to asymptotic value of the potential at infinity.}
energy that separates two different classes of solutions.
Negative-energy solutions are \emph{localized}, i.e. they exhibit bound-state behavior far away from the scatterer.
These are bound states and, by means of their normalizability, they are associated to \emph{discrete}, well-defined energies.
The bound states can be of course labelled by quantum numbers that are defined by the properties of the scatterer, namely the
symmetries of the potential $\op V$.
In a computational discretization of the Schr\"odinger equation, such states can be in principle expressed with arbitrary precision,
provided that the numerical treatment offers the adequate level of completeness.
Being discrete and well-identified, the bound-state energies are a property of the molecule and can be associated to \emph{observable}
quantities. The values of their energies can be compared among different treatments and it is in principle possible to
provide reference results on them.

@@

\begin{figure}[ht]
\includegraphics[scale=0.6]{Fig4_CO_dos_rev1.eps} %{Fig4_CO_dos.eps}
\caption{(Color online) DOS of $CO$ excitation energies $\Omega_a$. The two curves correspond to two computational domain of different sizes. The inset contains the sole contribution of localized excitations. The first IP threshold is depicted as the vertical dashed line.}
\label{CO_exc}
\end{figure}
\begin{figure}[ht]
\centering
\subfloat[Excitation landscape]
{\label{c6h6_excLand}
\includegraphics[scale=0.56]{Fig5a_C6H6_excLandscape.eps}} \\
\centering
\subfloat[DOS of Excitation energies]
{\label{C6H6_exc}
\includegraphics[scale=0.58]{Fig5b_C6H6_dos_rev1.eps}} %{Fig5b_C6H6_dos.eps}
\caption{(Color online) Benzene molecule. (a): Excitation Landscape: the excitations have been separated in discrete and continuum sector following the criteria of Sec.~\ref{ExcitationLandscape}. (b): DOS of excitation energies $\Omega_a$. The first IP threshold is depicted as the vertical dashed line.}
\end{figure}

@@

We have analyzed the fullfillment of Eq.~\eqref{sminus2} for molecules of various
sizes and symmetries.
In Fig.~\ref{co_AlphaExc} we show explicitly such a comparison in the case of the $CO$ molecule.
We notice that a considerable number of pseudo-continuum excitations is required
to find the correct result; the sector of the discrete excitations, even though
parametrised by localized, observable (in the sense of reproducible) excited states,
appears to be largely insufficient to express the reference fluctuation state. Its contribution is even negligible w.r.t.
the one coming from the continuum sector.

\begin{figure}
\includegraphics[scale=0.6]{Fig6_CO_statPolvsExc.eps}
\caption{\label{co_AlphaExc}(Color online) Convergence of the statical polarizability $\alpha_{zz}$ of the $CO$ molecule, in a box of $11.5$ \AA,
w.r.t the excitations considered in the integral of Eq.~\eqref{sminus2}. The reference value obtained from Eq.~\eqref{staticalpha} is represented as a
horizontal red dashed line. The value of the IP energy is indicated by a vertical black dashed line. For each excitation energy, the values of the
oscillator strengths $w^a_{zz}$ of Eq.~\eqref{sminus2} are also plotted. }
\end{figure}

@@

On the contrary, computational treatments that are able to express delocalized states may be adeguate to describe more efficiently the excitations DOS,
especially due to the capability of the basis to capture oscillatory behavior. Like in localized basis sets, excitations below threshold can be
precisely calculated with the same paradigm of traditional GS calculations. However, in this computational setup, to correctly express the linear susceptibility below IP -- more specifically the real part of the fluctuation state -- care should be taken in considering enough delocalized excitations
to guarantee fulfillment of sum-rules like Eq.~\eqref{sminus2}.

@@

\subsection{$CO$}
The molecule is oriented along the $z$ axis and the $z$-dimension of the box is equal to 11.5, 15.1 and 16.6 \AA.
Bond length is 1.235\AA, stretched wrt the equilibrium LDA position such as to increase the value of the polarizability.
280 unoccupied states have been computed in each setup, which span an energy range of the empty KS orbitals of 30.7, 19.3 and 15.4 eV respectively.
Such energy values should not be confused with the excitations energy ranges reported in Figs.~\ref{co_spectrum} and \ref{CO_exc}.
Once again, such choices of unoccupied states ensures convergence in of excitations spectrum in the presented range.

>>>>

From the above expressions we can see that, for molecular systems,the value $\epsilon=0$
represents a \emph{threshold}~\footnote{The threshold value in general should correspond to asymptotic value of the potential at infinity.}
energy that separates two different classes of solutions.
Negative-energy solutions are \emph{localized}, i.e. they exhibit bound-state behavior far away from the scatterer.
These are bound states and, by means of their normalizability, they are associated to \emph{discrete}, well-defined energies.
The bound states can be of course labeled by quantum numbers that are defined by the properties of the scatterer, namely the
symmetries of the potential $\op V$.
In a computational discretization of the Schr\"odinger equation, such states can be in principle expressed with arbitrary precision,
provided that the numerical treatment offers the adequate level of completeness.
Being discrete and well-identified, the bound-state energies are a property of the molecule and can be associated to \emph{observable}
quantities. The values of their energies can be compared among different treatments and it is in principle possible to
provide reference results on them.

@@

\begin{figure}[ht]
\includegraphics[scale=0.6]{Fig4_CO_dos_rev1.eps}
\caption{(Color online) DOS of $CO$ excitation energies $\Omega_a$. The two curves correspond to two computational domain of different sizes. The inset contains the sole contribution of localized excitations. The first IP threshold is depicted as the vertical dashed line.}
\label{CO_exc}
\end{figure}
\begin{figure}[ht]
\centering
\subfloat[Excitation landscape]
{\label{c6h6_excLand}
\includegraphics[scale=0.56]{Fig5a_C6H6_excLandscape.eps}} \\
\centering
\subfloat[DOS of Excitation energies]
{\label{C6H6_exc}
\includegraphics[scale=0.58]{Fig5b_C6H6_dos_rev1.eps}}
\caption{(Color online) Benzene molecule. (a): Excitation Landscape: the excitations have been separated in discrete and continuum sector following the criteria of Sec.~\ref{ExcitationLandscape}. (b): DOS of excitation energies $\Omega_a$. The first IP threshold is depicted as the vertical dashed line.}
\end{figure}

@@

We have analyzed the fulfillment of Eq.~\eqref{sminus2} for molecules of various
sizes and symmetries.
In Fig.~\ref{co_AlphaExc} we show explicitly such a comparison for the $CO$ and $C_{60}$ molecules.
We notice that a considerable number of pseudo-continuum excitations is required
to find the correct result; the sector of the discrete excitations, even though
parametrized by localized, observable (in the sense of reproducible) excited states,
appears to be largely insufficient to express the reference fluctuation state. Its contribution is even negligible w.r.t.
the one coming from the continuum sector.

% \begin{figure}
% \includegraphics[scale=0.6]{Fig6_CO_statPolvsExc.eps}
% \caption{\label{co_AlphaExc}(Color online) Convergence of the statical polarizability $\alpha_{zz}$ of the $CO$ molecule, in a box of $11.5$ \AA,
% w.r.t the excitations considered in the integral of Eq.~\eqref{sminus2}. The reference value obtained from Eq.~\eqref{staticalpha} is represented as a
% horizontal red dashed line. The value of the IP energy is indicated by a vertical black dashed line. For each excitation energy, the values of the
% oscillator strengths $w^a_{zz}$ of Eq.~\eqref{sminus2} are also plotted. }
% \end{figure}

\begin{figure}
\centering
\subfloat[$CO$]
{\label{CO_alphavsExc}
\includegraphics[scale=0.6]{Fig6a_CO_statPolvsExc.eps}} \\
\centering
\subfloat[$C_{60}$]
{\label{C60_alphavsExc}
\includegraphics[scale=0.2]{Fig6b_C60_statPolvsExc.eps}}
\caption{\label{co_AlphaExc}(Color online) Convergence of the statical polarizability $\alpha_{zz}$ of the $CO$ (panel a) and $C_{60}$ (panel b) molecules,
w.r.t the excitations considered in the integral of Eq.~\eqref{sminus2}. The reference value obtained from Eq.~\eqref{staticalpha} is represented as a
horizontal red dashed line. The value of the IP energy is indicated by a vertical black dashed line. For each excitation energy, the values of the
oscillator strengths $w^a_{zz}$ of Eq.~\eqref{sminus2} are also plotted. }
\end{figure}

@@

On the contrary, computational treatments that are able to express delocalized states may be adequate to describe more efficiently the excitations DOS,
especially due to the capability of the basis to capture oscillatory behavior. Like in localized basis sets, excitations below threshold can be
precisely calculated with the same paradigm of traditional GS calculations. However, in this computational setup, to correctly express the linear susceptibility below IP -- more specifically the real part of the fluctuation state -- care should be taken in considering enough delocalized excitations
to guarantee fulfillment of sum-rules like Eq.~\eqref{sminus2}.

@@

All the details of the calculations, the \textsl{Jupyter} notebooks of the data analysis, including the supplementary information, may be found at the URL~\footnote{\texttt{https://github.com/luigigenovese/LR-nb/blob/master/ POLARIZABILITY/nb\_paper/LR\_Analysis.ipynb}, or \texttt{https://gitlab.com/luigigenovese/ localization-of-excitations/ blob/master/LR\_Analysis.ipynb}}

\subsection{$CO$}
The molecule is oriented along the $z$ axis and the $z$-dimension of the box is equal to 11.5, 15.1 and 16.6 \AA.
Bond length is 1.235\AA, stretched w.r.t. the equilibrium LDA position such as to increase the value of the polarizability.
280 unoccupied states have been computed in each setup, which span an energy range of the empty KS orbitals of 30.7, 19.3 and 15.4 eV respectively.
These energy values should not be confused with the excitations energy ranges reported in Figs.~\ref{co_spectrum} and \ref{CO_exc}.
Once again, such choices of unoccupied states ensures convergence of the excitation spectrum in the presented range.
